---
title: "datasauRus"
author: "Kasia Banas"
date: "2024-09-26"
output: html_document
---

Install and load the datasauRus library:

```{r}
# install.packages("datasauRus")
library(datasauRus)
```

Load the tidyverse: 
```{r}
library(tidyverse)
```

`datasaurus_dozen` is a dataset demonstrating the utility of visualization. 
These 12 datasets are equal in standard measures: mean, standard deviation, 
and Pearson's correlation. But they look very different
Run the chunk below to get more info about this dataset:

```{r}
?datasaurus_dozen
```

See what datasets are in the datasaurus_dozen:

```{r}
datasaurus_dozen %>% 
  distinct(dataset) 
head(datasaurus_dozen)
```

## Exercise 1
Pick 2 datasets and plot them next to one another
(I'd recommend scatterplot, use facets if you like)

```{r}
star_dots <- datasaurus_dozen %>% 
  filter(dataset == "star" | dataset == "dino") 

star_dots %>%
  ggplot(aes(x = x, y = y, colour = dataset))+
  geom_point()+
  facet_wrap(~dataset)
```


## Exercise 2
For these same two datasets, find the mean and standard deviation
for the x and y variable in each. What did you find? 
```{r}
dinostat <- datasaurus_dozen %>% 
  filter(dataset == "dino")
sd(dinostat$x, na.rm = FALSE)
sd(dinostat$y, na.rm = FALSE)
#mean: x=54.26, y=47.83
#SD: x= 16.765, y=26.935

starstat <- datasaurus_dozen %>% 
  filter(dataset == "star")
sd(starstat$x, na.rm = FALSE)
sd(starstat$y, na.rm = FALSE)
#mean: x=54.27, y=47.84
#SD: x=16.77 , y=26.93

datasaurus_dozen %>% 
  filter(dataset == "dino" | dataset =="star") %>% 
  group_by(dataset) %>% 
  summarise(mean_x= mean(x), sd_x = (sd(x)), mean_y = mean(y), sd_y = sd(y))
```

